---
title: "MA 4780 Homework 2"
author: "Benjamin Hendrick"
date: "February 10, 2016"
output:
  pdf_document:
    fig_caption: yes
    fig_height: 3
    fig_width: 4
---

# Problem 3.2

\begin{align*}
\bar{Y} &= \big[ \mu + \frac{1}{n} \sum^{n}_{t=1} (e_{t} - e_{t-1}) \big] \\
        &= \mu + \frac{1}{n}(e_{n}-e_{0})
\end{align*}

Therefore, 

\begin{align*}
Var(\bar{Y}) &= \frac{1}{n^{2}} Var(e_{n}-e_{0}) \\
             &= \frac{2}{n^{2}} \sigma^{2}_{e}
\end{align*}

# Problem 3.3

$$\sum^{n}_{t=1} (e_{t} + e_{t-1}) = e_{n} + e_{0} + 2\sum^{n-1}_{t=1} e_{t}$$

Therefore,

\begin{align*}
Var(\bar{Y}) &= \frac{1}{n^{2}} [\sigma^{2}_{e} + \sigma^{2}_{e} + 4(n-1)\sigma^{2}_{e}] \\
             &= \frac{2(2n-1)}{n^2} \sigma^{2}_{e}
\end{align*}

# Problem 3.7

Load the data.

```{r, results='hide'}
library(TSA)
data("winnebago")
```

## Part A

Figure \ref{fig:one} shows the time series plot of the data `winnebago`.

```{r fig.cap="Time series plot of monthly unit sales of recreational vehicles from Winnebago, Inc. \\label{fig:one}"}
plot(winnebago,
     type="o",
     ylab="Unit Sales",
     main="Monthly Unit Sales over Time")
model <- lm(winnebago~time(winnebago))
abline(model)
```

## Part B

The fitted regression line fits well to the time series data. The line data is not heteroscedastic, but the line still fits well enough. Figure \ref{fig:two} shows the standardized residuals from the fit as a time series. The residuals appear to be centered around mean zero.

```{r fig.cap="Plot of standardized residuals from the fit as a time series. \\label{fig:two}"}
plot(rstudent(model),
     type = "o",
     ylab = "Standard Residuals",
     main = "Monthly Unit Sales")
```

## Part C

Figure \ref{fig:three} time series plot of the monthly unit sales with a log scale. 

```{r fig.cap = "Time series plot of monthly unit sales of recreational vehicles from Winnebago, Inc. under a log scale \\label{fig:three}"}
winnLog <- log(winnebago)
plot(winnLog, 
     type="o",
     ylab = "Unit Sales (Log Scale)", 
     main = "Monthly Unit Sales")
logModel <- lm(winnLog~time(winnLog))
abline(logModel)
```

## Part D

The least squares fit on the logged data fits very well. The data with the log transformation seem more heteroscedastic, yet just as linear. Figure \ref{fig:four} shows the standardized residuals of the logged data. The residuals appear to be centered around mean zero.

```{r fig.cap = "Plot of standardized residuals from the fit as a time series with log scale. \\label{fig:four}"}
plot(rstudent(model),
     type = "o",
     ylab = "Standard Residuals",
     main = "Monthly Unit Sales")
```

## Part E

```{r}
month <- season(winnebago)
winCombo <- lm(winnLog~month + time(winnLog))
summary(winCombo)
```

Many of the month coefficients are statistically significant. All months but October, November, and December are not statistically significant. This is because their p-values were too high (greeter than 0.10). 

## Part F

Figure \ref{fig:five} shows the standardized residuals from the combined fit. The residuals appear to be centered around mean zero.

```{r fig.cap="Plot of standardized residuals from the combined fit as a time series. \\label{fig:five}"}
plot(rstudent(winCombo), 
     type = "o", 
     ylab = "Standard Residuals", 
     main ="Monthly Unit Sales")
```


# Problem 3.8

Load the data.

```{r}
data(retail)
```

## Part A

Figure \ref{fig:six} shows the time series data with seasonal markers. There is clear seasonality where December has the highest sales. The overall sales is positively increasing.

```{r fig.cap= "The time series plot of the retail data with seasonal symbols. \\label{fig:six}"}
plot(retail,
     type = "l",
     ylab = "Billions of Pounds",
     main = "Retail Sales over Time")
points(y=retail,
       x=time(retail), 
       pch=as.vector(season(retail)))
```


## Part B

The linear model has many important coefficients. The `summary` of the model shows that every month is statistically significant. This is because they have small p-values (less than 0.05).

```{r}
retail.lm <- lm(retail~season(retail) + time(retail))
summary(retail.lm)
```


## Part C

Figure \ref{fig:seven} shows the standardized residuals. The residuals appear symmetric over time, implying that they are centered around mean zero.

```{r fig.cap= "Plot of standardized residuals from the model in Part B \\label{fig:seven}"}
plot(rstudent(retail.lm),
     type = "l",
     ylab = "Standardized Residuals",
     main = "Retail Sales")
points(y=rstudent(retail.lm),
       x=as.vector(time(rstudent(retail.lm))), 
       pch=as.vector(season(retail)))
```


# Problem 3.9

Load the data.

```{r}
data(prescrip)
```

## Part A

Figure \ref{fig:eight} shows the time series plot with seasonal markings. There is an apparent seasonal trend during the summer. The overall trend is positively increasing.

```{r fig.cap= "Time series plot of the monthly prescription costs in the United States \\label{fig:eight}"}
plot(prescrip,
     type = "l",
     ylab = "Presciption Costs",
     main = "Presciption Costs over Time")
points(y = prescrip,
       x = time(prescrip),
       pch = as.vector(season(prescrip)))
```

## Part B

Figure \ref{fig:nine} shows the time series plot of the percentage changes in prescription costs with seasonal markers.

```{r fig.cap="Time series of the sequence of month-to-month percentage changes in prescription costs. \\label{fig:nine}"}
percPrecrip <- na.omit(100*(prescrip - zlag(prescrip))/zlag(prescrip))
plot(percPrecrip,
     type = "l",
     ylab = "Percent Change of Prescription Costs",
     main = "Percent Change of Prescription Costs over Time")
points(y = percPrecrip,
       x = time(percPrecrip),
       pch = as.vector(season(percPrecrip)))
```


## Part C

```{r}
prescrip.lm <- lm(percPrecrip ~ harmonic(percPrecrip))
summary(prescrip.lm)
```

The `summary` of the data shows that the harmonic cosine coefficient is not statistically significant because the p-value is greater than 0.05. If our alpha values is set to 0.10, then every coefficient would be statistically significant.

## Part D

Figure \ref{fig:ten} shows the standardized residuals of the cosine model. The residuals appear to be centered around mean zero and are fairly random.

```{r fig.cap="Plot of standardized residuals of the cosine model. \\label{fig:ten}"}
plot(rstudent(prescrip.lm),
     type = "l",
     ylab = "Standardized Residuals",
     main = "Presciption Costs")
points(y=rstudent(prescrip.lm),
       x=as.vector(time(rstudent(prescrip.lm))), 
       pch=as.vector(season(prescrip)))
```

# Problem 3.13

Load the data.

```{r}
data("retail")
```

## Part A

```{r}
winnebago.lm <- lm(winnebago~season(log(winnebago)) + time(log(winnebago)))
```

## Part B

```{r}
runs(rstudent(winnebago.lm))
```

The `runs` test suggest a lack of independence in the error terms in the model.

## Part C

Figure \ref{fig:eleven} shows the ACF plot of the combined Winnebago model. The residuals of lags one, two, three, and five suggest significant positive auto-correlation.

```{r fig.cap="ACF plot of combined Winnebago model residuals \\label{fig:eleven}"}
acf(rstudent(winnebago.lm))
```

## Part D

Figure \ref{fig:twelve} suggests that the distribution of the standardized residuals of the combined Winnebago model is right skewed. Figure \ref{fig:thirteen} suggests the same.

```{r fig.cap="Histogram of combined Winnebago model residuals. \\label{fig:twelve}"}
hist(rstudent(winnebago.lm),
     xlab = "Standardized Residuals",
     main = "Historgram of Standardized Residuals")
```

```{r fig.cap="QQ plot of combined Winnebago model residuals.\\label{fig:thirteen}"}
qqnorm(rstudent(winnebago.lm))
qqline(rstudent(winnebago.lm))
```


# Problem 3.14

Load the data.

```{r}
data("retail")
```

## Part A

```{r}
retail.lm <- lm(retail~season(retail) + time(retail))
```

## Part B

```{r}
runs(rstudent(retail.lm))
```

The `runs` test suggest a lack of independence in the error terms in the model.

## Part C

```{r fig.cap="ACF plot for combined retail model.\\label{fig:fourteen}"}
acf(rstudent(retail.lm))
```

Figure \ref{fig:fourteen} suggest significant positive auto-correlation at lags one, eleven, twelve, thirteen, and twenty-four.

## Part D

Figure \ref{fig:fifteen} suggests that the distribution of the standardized residuals of the combined retail model is relatively normal. Figure \ref{fig:sixteen} suggests the same.


```{r fig.cap="Histogram of standardized residuals of combined retail model. \\label{fig:fifteen}"}
hist(rstudent(retail.lm),
     xlab = "Standardized Residuals",
     main = "Historgram of Standardized Residuals")
```

```{r fig.cap="QQ plot of standardized residuals of combined retail model.\\label{fig:sixteen}"}
qqnorm(rstudent(retail.lm))
qqline(rstudent(retail.lm))
```

# Problem 3.15
Load the data.

```{r}
data("prescrip")
```

## Part A
```{r}
percPrecrip <- na.omit(100*(prescrip - zlag(prescrip))/zlag(prescrip))
prescrip.lm <- lm(percPrecrip ~ harmonic(percPrecrip))
```

## Part B
```{r}
runs(rstudent(prescrip.lm))
```

The `runs` test suggest a lack of independence in the error terms in the model.

## Part C

Figure \ref{fig:seventeen} suggests significant negative auto-correlation at lags one and thirteen.

```{r fig.cap="ACF plot for the precription model.\\label{fig:seventeen}"}
acf(rstudent(prescrip.lm))
```


## Part D

Figure \ref{fig:eighteen} suggests that the distribution of the standardized residuals of the combined retail model is relatively normal. Figure \ref{fig:nineteen} suggests the same.

```{r fig.cap="Histogram of standardized residuals of combined retail model. \\label{fig:eighteen}"}
hist(rstudent(prescrip.lm),
     xlab = "Standardized Residuals",
     main = "Historgram of Standardized Residuals")
```

```{r fig.cap="QQ plot of standardized residuals of combined retail model.\\label{fig:nineteen}"}
qqnorm(rstudent(prescrip.lm))
qqline(rstudent(prescrip.lm))
```